\vspace{-0.5in}{\bfseries\scshape Transfer Learning to train Vision Transformers with limited resources } \\ \vspace{1ex} \normalsize{Bachelor's Project Thesis}\vspace{-2ex}

### 2022-04-23 writing/tex/draft_introduction.tex
Almost there :-)

The self-attention mechanism plays a central role in ViTs, and is worth elaborating on. Each head of a self-attention layer contains 3 learned projection matrices, which are used to produce $key$, $value$, and $query$ vectors for every position in the input sequence. The output for a single position is a weighted sum of all $value$ vectors. Weights are obtained by combining this position's $query$ with all $key$ vectors\footnote{In principle by taking softmax over all query-key dot products.}. As such, an attention head decides itself how much of each position should be incorporated in subsequent representations.

@@

In practice, many TL techniques exist. The example given in section \ref{vits} can be described as \textit{fine-tuning} TL. Here, a pre-trained model is used as a starting point, and trained fully on the target task -- possibly with different hyper-parameters, and often with a new final layer that fits the task. \citeauthor{yosinski2014transferable} (\citeyear{yosinski2014transferable}) studied transferability properties of CNNs pre-trained on ImageNet, and suggested that \textit{fine-tuning} results in improved performance and generalizability compared to training from scratch. Moreover, while transferability decreased as source and target task became more dissimilar, utilizing TL still seemed to be preferred.

@@

A different TL technique is called \textit{off-the-shelf} (OTS) TL. It is similar to \textit{fine-tuning}, except that all but the final layer is kept frozen during training. \citeauthor{sharif2014cnn} (\citeyear{sharif2014cnn}) found that features extracted by a CNN pre-trained on ImageNet were general enough for OTS TL to work. More specifically, these OTS CNNs showed better performance than the (fully trained non-CNN) state of the art at the time. Besides reducing computational cost of training, a benefit of OTS TL is that it can prevent overfitting in cases where \textit{fine-tuning} would not \citep{yosinski2014transferable}.

@@

\subsection{The current study and related work}
TODO

>>>>

The self-attention mechanism plays a central role in ViTs, and is worth elaborating on. Each head of a self-attention layer contains 3 learned projection matrices, which are used to produce $key$, $value$, and $query$ vectors for every position in the input sequence. The output for a single position is a weighted sum of all $value$ vectors. Weights are obtained by matching this position's $query$ with all $key$ vectors\footnote{In principle by taking softmax over all query-key inner products.}. As such, an attention head decides itself how much of each position should be incorporated in subsequent representations.

@@

In practice, many TL techniques exist. The example given in section \ref{vits} can be described as \textit{fine-tuning} TL. Here, a pre-trained model is used as a starting point, and then trained on the target task -- possibly with different hyper-parameters, and often with a new final layer that fits the task. \citeauthor{yosinski2014transferable} (\citeyear{yosinski2014transferable}) studied transferability properties of CNNs pre-trained on ImageNet, and suggested that fine-tuning results in improved performance and generalizability compared to training from scratch. Moreover, while transferability decreased as source and target task became more dissimilar, utilizing TL still seemed to be preferred.

@@

A different TL technique is called \textit{off-the-shelf} (OTS) TL. It is similar to fine-tuning, except that all but the final layer is kept frozen during training. \citeauthor{sharif2014cnn} (\citeyear{sharif2014cnn}) found that features extracted by a CNN pre-trained on ImageNet were general enough for OTS TL to work. More specifically, these OTS CNNs showed better performance than the (fully trained non-CNN) state of the art at the time. Besides reducing computational cost of training, a benefit of OTS TL is that it can prevent overfitting in cases where fine-tuning would not \citep{yosinski2014transferable}.

@@

\subsection{Related work and the current study}
As mentioned, the paper that first introduced ViTs already demonstrated they benefit from TL \citep{dosovitskiy2020image}, and more has been discovered about ViTs' TL properties since then. \citeauthor{matsoukas2021time} (\citeyear{matsoukas2021time}) investigated whether TL allowed ViTs to replace CNNs in Medical Imaging. This domain is characterized by small datasets, so to overcome the inductive bias problem described earlier, TL is imperative. It was found that ViTs benefit more from fine-tuning TL than CNNs do, and that ViTs pre-trained on ImageNet perform on par CNNs.

What \citeauthor{matsoukas2021time} did not investigate, is OTS performance of ViTs. This was, however, done in \citeauthor{zhou2021convnets} (\citeyear{zhou2021convnets}), which found that ViTs pre-trained on ImageNet had worse OTS performance than their CNN counterparts. An exception was the \textit{Swin-B} architecture, which was the best performing model on all datasets, and uses \textit{(shifted-)window-based} self-attention methods, to make Transformers better suited for CV tasks \citep{liu2021swin}.
% Not into detail about ots and other techniques

Another ``shortcomming'' of \citeauthor{matsoukas2021time} is that it merely examined whether ViTs can use TL to outperform CNNs in a domain that happens to be characterized by limited-sized datasets, but that it did not actively focus on the extent to which ViTs can be made useful in limited data/compute environments. More related to this is \citeauthor{touvron2021training} (\citeyear{touvron2021training}), which proposed transformer-specific distillation-based learning techniques to make ViTs applicable for smaller datasets. This is not specifically related to TL as defined in this paper, however.
% matsoukas2021time not specifically focussed on seeing to what degree TL allows ViTs to dominate CNNs in limited compute/data environments
